Formalized Linear Programming

### 2019-09-23 TeX/paper.tex
minor changes

\begin{document}
\bibliographystyle{plain}
\lstset{language=lean}
\lstset{basicstyle=\normalsize}
\maketitle
\begin{abstract}
  Sentence about applications of simplex. Most simplex solvers are non-verified implementations and are not guaranteed to be correct. They are usually written in languages like C++ which are difficult to certify. One verified implementation exists, but it is not intended to be executable in practice. We implement a formally verified executable simplex solver in the Lean theorem prover. Our implementation uses Bland's rule to select a pivot, and a pivot function stated without any matrix multiplication in contrast to some book proofs on the simplex algorithm.
\end{abstract}

@@

The type \lstinline|cvec (m + n)| is notation for $m + n \times 1$ matrices. \lstinline|x i 0| is the notation for the $i$th element of this vector.

@@

@@

The partial equivalence corresponding to $R$ is given by the vector $r$,

@@

\begin{lstlisting}
inductive rel : tableau m n → tableau m n → Prop
| pivot : ∀ {T}, feasible T → ∀ {r c}, c ∈ pivot_col T obj →
  r ∈ pivot_row T obj c → rel (T.pivot r c) T
| trans_pivot : ∀ {T₁ T₂ r c}, rel T₁ T₂ → c ∈ pivot_col T₁ obj →
  r ∈ pivot_row T₁ obj c → rel (T₁.pivot r c) T₂
\end{lstlisting}

\color{red}Include description of pivot\_col and pivot\_row functions somewhere. Explain $\in$ notation for option. \color{black}

Two tableau \lstinline|T| and \lstinline|T'| are related if either \lstinline|T' = T.pivot r c| where \lstinline|r| and \lstinline|c| are the pivot row and column selected according to Bland's rule, or if there exists another tableau \lstinline|T₁| such that \lstinline|rel obj T₁ T| and \lstinline|T' = T₁.pivot r c|.

>>>>

\begin{document}
\bibliographystyle{plain}
\lstset{language=lean}
\lstset{basicstyle=\normalsize}
\maketitle
\begin{abstract}
  Sentence about applications of simplex. Most simplex solvers are non-verified implementations and are not guaranteed to be correct. They are usually written in languages like C++ which are difficult to certify. One verified implementation exists, but it is not intended to be executable in practice. We implement a formally verified executable simplex solver in the Lean theorem prover. Our implementation uses Bland's rule to select a pivot, and a pivot function stated without any matrix multiplication in contrast to some book proofs on the simplex algorithm. This research demonstrates the practicality of Lean as a language for fast verified algorithms.
\end{abstract}

@@

The type \lstinline|cvec (m + n)| is notation for $m + n \times 1$ matrices. \lstinline|x i 0| is the notation for the $i$th element of the vector  \lstinline|x|.

@@

\begin{lstlisting}
@[simp] lemma sol_set_simplex (T : tableau m n) (hT : feasible T) (w : tableau m n → bool)
  (obj : fin m) : (T.simplex w obj hT).1.sol_set = T.sol_set

lemma termination_eq_while_iff {T : tableau m n} {hT : feasible T} {w : tableau m n → bool}
  {obj : fin m} : (T.simplex w obj hT).2 = while ↔ w (T.simplex w obj hT).1 = ff

lemma termination_eq_optimal_iff {T : tableau m n} {hT : feasible T}
  {w : tableau m n → bool} {obj : fin m} : (T.simplex w obj hT).2 = optimal ↔
  w (T.simplex w obj hT).1 = tt ∧
  is_optimal T ((T.simplex w obj hT).1.of_col 0) (T.to_partition.rowg obj)

lemma termination_eq_unbounded_iff {T : tableau m n} {hT : feasible T}
  {w : tableau m n → bool} {obj : fin m} {c : fin n}: (T.simplex w obj hT).2 = unbounded c ↔
  w (T.simplex w obj hT).1 = tt ∧ is_unbounded_above T (T.to_partition.rowg obj)
  ∧ c ∈ pivot_col (T.simplex w obj hT).1 obj
\end{lstlisting}

@@

The partial equivalence corresponding to $R$ is given by the vector $r$, mapping $i$ to $r_i$. This is an injective function and therefore a bijection with its own image. The matrix $R$ is given by
\begin{equation}
R_{ij} =
\begin{cases}
1 & \text{if } j = r_i \\
0 & \text{otherwise}
\end{cases}
\end{equation}

@@

\begin{lstlisting}
inductive rel : tableau m n → tableau m n → Prop
| pivot : ∀ {T}, feasible T → ∀ {i j}, pivot_col T obj = some j →
  pivot_row T obj j = some i → rel (T.pivot i j) T
| trans_pivot : ∀ {T₁ T₂ i j}, rel T₁ T₂ → pivot_col T₁ obj = some j →
  pivot_row T₁ obj j = some i → rel (T₁.pivot i j) T₂
\end{lstlisting}

Two tableau \lstinline|T| and \lstinline|T'| are related if either \lstinline|T' = T.pivot i j| where \lstinline|i| and \lstinline|j| are the pivot row and column selected according to Bland's rule, or if there exists another tableau \lstinline|T₁| such that \lstinline|rel obj T₁ T| and \lstinline|T' = T₁.pivot i j|.
